A Horrocks-type theorem for even orthogonal $K_2$

### 2019-09-24 dl-paper.tex
Minor

For an additive subgroup $A\subseteq R$ and $\alpha \in \Phi$ we denote by $X_\alpha(A)$ the corresponding {\it root subgroup} of $\St(\Phi, R)$, i.\,e. the subgroup generated by all $x_\alpha(a)$, $a \in A$.

@@

\subsection{Computation of the kernel of the map of evaluation at $0$.}
For the rest of this subsection $A$ denotes a local ring with maximal ideal $M$.
The aim of this subsection is to describe a generating set for the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. We denote this kernel by $K(A[X], M[X])$.

It is obvious that $K(A[X], M[X])$ contains the subgroup $\overline{\St}(\Phi, A[X], XM[X])$. It turns out that, although being larger than this subgroup, $K(A[X], M[X])$ contains very few extra generators, which can all be explicitly described (see~\cref{Kgen} and the corollary that follows it).

@@

 Now each factor $z_{\alpha_i}(f_i(0), \xi_i(X))$ can be written as follows:
 \[z_{\alpha_i}(f_i(0), \xi_i(0))^{x_{-\alpha_i}(X\xi'_i(X))} = [x_{-\alpha_i}(-X\xi'_i(X)), z_{\alpha_i}(f_i(0), \xi_i(0))] \cdot z_{\alpha_i}(f_i(0), \xi_i(0)).\]
 Since the subgroup $C_0 = \left[\St(\Phi, XA[X]), \overline{\St}(\Phi, A, M)\right]$ is normalized by $\overline{\St}(\Phi, A, M)$ 
 (this follows e.\,g. from the formula $[g, h]^{h_1} = [h_1^{-1}, g][g, h_1^{-1}h]$) we conclude that $g_1(X)$ is congruent to $\prod_i z_{\alpha_i}(f_i(0), \xi_i(0)) = g(0) = 1$ modulo $C_0$,
 which implies the assertion. \qedhere \end{proof}

@@

\begin{prop} \label{Kgen} The subgroup $K(A[X], M[X])$ is generated as an abstract group by the subgroup $\overline{\St}(\Phi, A[X], XM[X])$ and
 the elements $[x_\alpha(m), x_{-\alpha}(X\xi)]$, $m \in M$, $\xi \in A[X]$, $\alpha \in \Phi$. \end{prop}
\begin{proof} Notice that it follows from~\eqref{eq:conj-h-x} $\overline{H}(\Phi, A, M)$ normalizes both $\St(\Phi, XA[X])$ and $\St(\Phi, M)$ and, moreover, that $[\overline{H}(\Phi, A, M),\ \St(\Phi, XA[X])] \subseteq \overline{\St}(\Phi, A[X], XM[X])$.

Denote by $C_1$ the commutator subgroup $[\St(\Phi, XA[X]), \St(\Phi, M)]$.
It is clear that for $g \in \St(\Phi, XA[X])$, $h \in \overline{H}(\Phi, A, M)$, $u^+ \in U(\Phi^+, M)$, $u^- \in U(\Phi^-, M)$ holds:
\[ [g, h u^+ u^-] = [g, h] \cdot [{}^{h}\!g, {}^{h}\!(u^+u^-)] \in \St(\Phi, A[X], XM[X]) \cdot C_1.\]
Since $C_0$ is generated by the above commutators and $\overline{\St}(\Phi, A[X], XM[X])$ is a normal subgroup of $\St(\Phi, A[X])$
we obtain that $C_0 \subseteq \overline{\St}(\Phi, A[X], XM[X]) \cdot C_1$ and consequently that
$K(A[X], M[X]) = \overline{\St}(\Phi, A[X], XM[X]) \cdot C_1.$

>>>>

For an additive subgroup $A\subseteq R$ and $\alpha \in \Phi$ we denote by $X_\alpha(A)$ the corresponding {\it root subgroup} of $\St(\Phi, R)$, i.\,e. the subgroup generated by all $x_\alpha(a)$, $a \in A$.
We denote by $U(\Phi^+, A)$ (resp. $U(\Phi^-, A)$) the subgroup of $\St(\Phi, R)$ generated by root subgroups $X_\alpha(A)$ corresponding to all positive (resp. negative) roots of $\Phi$.

@@

\subsection{Computation of the kernel of the map of evaluation at $0$.}
Let $A$ be a local ring with maximal ideal $M$.
The aim of this subsection is to describe a generating set for the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. We denote this kernel by $K(A[X], M[X])$.

It is obvious that $K(A[X], M[X])$ contains the subgroup $\overline{\St}(\Phi, A[X], XM[X])$.
It turns out that, although $K(A[X], M[X])$ is generally strictly larger than $\overline{\St}(\Phi, A[X], XM[X])$,
 it contains very few extra generators, which can all be explicitly described (see~\cref{Kgen} and the corollary that follows it).

@@

 Now each factor $z_{\alpha_i}(f_i(0), \xi_i(X))$ can be written as follows:
 \[z_{\alpha_i}(f_i(0), \xi_i(0))^{x_{-\alpha_i}(X\xi'_i(X))} = [x_{-\alpha_i}(-X\xi'_i(X)),\ z_{\alpha_i}(f_i(0), \xi_i(0))] \cdot z_{\alpha_i}(f_i(0), \xi_i(0)).\]
 Since the subgroup $C_0 = \left[\St(\Phi, XA[X]), \overline{\St}(\Phi, A, M)\right]$ is normalized by $\overline{\St}(\Phi, A, M)$ 
 (this follows e.\,g. from the formula $[g, h]^{h_1} = [h_1^{-1}, g][g, h_1^{-1}h]$) we conclude that $g_1(X)$ is congruent to $\prod_i z_{\alpha_i}(f_i(0), \xi_i(0)) = g(0) = 1$ modulo $C_0$,
 which implies the assertion. \qedhere \end{proof}

@@

\begin{prop} \label{Kgen} The subgroup $K(A[X], M[X])$ is generated as an abstract group by the subgroup $\overline{\St}(\Phi, A[X], XM[X])$ and
 the elements $[x_\alpha(m), x_{-\alpha}(X\xi)]$, $m \in M$, $\xi \in A[X]$, $\alpha \in \Phi$. \end{prop}
\begin{proof} From~\eqref{eq:conj-h-x} we obtain that $\overline{H}(\Phi, A, M)$ normalizes both $\St(\Phi, XA[X])$ and $\St(\Phi, M)$ and, moreover, that $[\overline{H}(\Phi, A, M),\ \St(\Phi, XA[X])] \subseteq \overline{\St}(\Phi, A[X], XM[X])$.

Denote by $C_1$ the commutator subgroup $[\St(\Phi, XA[X]),\ \St(\Phi, M)]$.
It is clear that for $g \in \St(\Phi, XA[X])$, $h \in \overline{H}(\Phi, A, M)$, $u^+ \in U(\Phi^+, M)$, $u^- \in U(\Phi^-, M)$ holds:
\[ [g,\ h u^+ u^-] = [g,\ h] \cdot [{}^{h}\!g,\ {}^{h}\!(u^+u^-)] \in \St(\Phi, A[X], XM[X]) \cdot C_1.\]
Since $C_0$ is generated by the above commutators and $\overline{\St}(\Phi, A[X], XM[X])$ is a normal subgroup of $\St(\Phi, A[X])$
we obtain that $C_0 \subseteq \overline{\St}(\Phi, A[X], XM[X]) \cdot C_1$ and consequently that
$K(A[X], M[X]) = \overline{\St}(\Phi, A[X], XM[X]) \cdot C_1.$
